---
title: DATA 606 Data Project Proposal
author:  Javier Pajuelo Bazan
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

Is CryoSleep predictive of a passenger was Transported=True?

### Cases 

**What are the cases, and how many are there?**
Each case represents a passenger.

8700 in the train set
4300 in the test set



### Data collection 

**Describe the method of data collection.**

Data is collected from Kaggle. It belongs to a competition and I believe it's an imaginary scenario.


### Type of study 

**What type of study is this (observational/experiment)?**

This is an observational study.


### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

Data was not self collected, but retrieved from Kaggle.

https://www.kaggle.com/competitions/spaceship-titanic/overview
Cite it.



### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

The response variable is the Transported boolean and it is numerical (0 or 1)



### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
library(tidyverse)
library(curl)
library(dplyr)                      

load_csv_from_url <- function(url_path)
{
  tmp <- tempfile()
  curl_download(url_path, tmp)
  read_csv(tmp)
}
 
approval_polls <- load_csv_from_url("https://raw.githubusercontent.com/fivethirtyeight/covid-19-polls/master/covid_approval_polls.csv")

```