---
title: "Project 3"
authors and collaborators: Cassandra Boylan, Trang Do, Victoria McEleney, Javier Pajuelo
  and Leticia Salazar
date: "`r Sys.Date()`"
output:
  html_document:
    theme: journal
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Project – Data Science Skills**


##### W. Edwards Deming said, “In God we trust, all others must bring data.” **Please use data to answer the question, “Which are the most valued data science skills?”** Consider your work as an exploration; there is not necessarily a “right answer.”

#### Objective: As a group our we worked on answering the question "Which are the most valued data science skills?" We looked into a dataset from [Data.World](https://data.world/jobspikr/10000-data-scientist-job-postings-from-the-usa/workspace/file?filename=data_scientist_united_states_job_postings_jobspikr.csv] and a downloaded HTML files from Indeed [https://www.indeed.com/jobs?l=New%20York%20State&vjk=ae3c34627b58db77) and from [Indeed.com](https://www.indeed.com/jobs?q=data%20scientist&start=). Once the data sets were aquired we tidied and transformed the dataset and performed visualization. Finally, we analyzed the datasets, which you'll see towards the end.

##### Collaboration tools: Our team collaborated virtually through Zoom meetings, Slack and GitHub. We also utilized R and MySQL as software tools.


#### We install our libraries
```{r}
library(tidyverse)
library(curl)
#install.packages("tm") # if not already installed
library(tm)

load_csv_from_url <- function(url_path)
{
  tmp <- tempfile()
  curl_download(url_path, tmp)
  read_csv(tmp)
}
```


#### Loading the data from GitHub
```{r}
jobspikr_url = 'https://raw.githubusercontent.com/quaere1verum/sps_public/master/data607-001/assignments/project3/data_scientist_united_states_job_postings_jobspikr.csv'

jobspikr_small_url = 'https://raw.githubusercontent.com/quaere1verum/sps_public/master/data607-001/assignments/project3/data_scientist_jobspikr_10152021_merged.csv'


jobspikr_data <- load_csv_from_url(jobspikr_small_url)
```


#### Tidying Data.World Data
```{r}
#Getting inferred_skills column from dataset
get_inferred_skills <- function(jobspikr_data)
{
  inferred_skills <- jobspikr_data[['inferred_skills']]
  inferred_skills <- strsplit(inferred_skills, "\\|")
  skills <- c()
  for(inferred_skill in inferred_skills){
    skills <-c(skills, inferred_skill)
  }
  # remove duplicates
  return(skills[!duplicated(skills)])
}

inferred_skills <- get_inferred_skills(jobspikr_data)
```


#### SQL Insertion Portion:

##### Assumptions:
* Windows Machine
* MySQL Workbench 8.0
  * there exists a user: 'root' &  server: 'localhost' with password=''. No password required.
* RStudio installed

###### Instructions:
* 1.  git clone *.git or retrieve raw files from [GitHub] (https://github.com/quaere1verum/sps_public)
* 2.  Start MySQL Workbench
  * a.) If your MySQL service does not start automatically when Windows starts, press Windows key and search for "Services", press enter. Search for MySQL80 and start the service.
  * b.) Open the file project3.sql and run it. 
* 3. Start RStudio
  * a.) Open and run the project3_code.r script
  

```{r}
library(digest)

tmp_frame <- jobspikr_data %>% mutate(companyid=unlist(lapply(company_name, function(x) {digest(x, algo="md5", serialize = F)})   ))

companies <- tmp_frame %>% select(companyid, company_name)

# Table ready for insertion
companies_table <- distinct(companies, companyid, company_name)
```


```{r}
#install.packages("DBI")
library(DBI)
#install.packages("RMySQL")
library(RMySQL)


# Trying tmp tables
dbWriteTable(con,"myTempTable", companies_table)
dbExecute(con,"insert into companies(companyid, company_name) select companyid, company_name from myTempTable")
dbExecute(con,"drop table if exists myTempTable")
dbExecute(con,"commit;")

# companies data is your data frame with the same schema as companies table defined under sql
insert_into_companies_table<- function(companies_data)
{
  dbWriteTable(con,"myTempTable", companies_data)
  dbExecute(con,"insert into companies(companyid, company_name) select companyid, company_name from myTempTable")
  dbExecute(con,"drop table if exists myTempTable")
  dbExecute(con,"commit;")
}

# same thing, schema needs to be the same
insert_into_skill_types_table <- function(skill_data)
{
  dbWriteTable(con,"myTempTable", skill_data)
  dbExecute(con,"insert into skill_types(skill_id, skill_name) select skill_id, skill_name from myTempTable")
  dbExecute(con,"drop table if exists myTempTable")
  dbExecute(con,"commit;")
  
}

# Skill Set should be in Database
test<-strsplit(jobspikr_data$inferred_skills, split = "\\|")
skillset <-data.frame(skill=character())
s <-length(test)

for (i in 1:s){
  for (j in 1:lengths(test[i])){
    rows<-data.frame(skill=test[[i]][j])
    skillset <-rbind(skillset,rows)
  }  
}

word_freq<- skillset %>% group_by(skill)%>%   summarise(wfreq=n()) 

# drop skill names that are NA doesn't make sense
word_freq <- word_freq %>% drop_na() 
tmp_frame <- word_freq %>% mutate(skill_id=unlist(lapply(skill, function(x) {digest(x, algo="md5", serialize = F)})), skill_name=skill)

# table ready for insertion
skill_data <- tmp_frame %>% select(skill_id, skill_name)
insert_into_skill_types_table(skill_data)
```


#### Retrieval of graphs based on frequency
```{r}
# Looks like job description has a lot of things mixed in it. Requirements, disclaimers, company culture descriptions
# character array 
job_description_data <- jobspikr_data[['job_description']]


data<- job_description_data
data <- jobspikr_data[['inferred_skills']]

# What happens if we pass data without | and instead spitted data?
inferred_skills <- jobspikr_data[['inferred_skills']]
inferred_skills <- strsplit(inferred_skills, "\\|")
data <- inferred_skills

# Put the data into a corpus for text processing
text_corpus <- (VectorSource(data))
text_corpus <- Corpus(text_corpus)
summary(text_corpus)


# To see the text and examine the corpus
text_corpus[[5]]$content
for (i in 1:5) print (text_corpus[[i]]$content)


##Tokenization: Split a text into single word terms called "unigrams" 
text_corpus_clean<-Boost_tokenizer(text_corpus)
#text_corpus_clean[[1]]$content


#Example in R: by using tm package
#Normalization: lowercase the words and remove punctuation and numbers
text_corpus_clean<-tm_map(text_corpus , content_transformer(tolower))
text_corpus_clean <- tm_map(text_corpus_clean, removePunctuation)
text_corpus_clean <- tm_map(text_corpus_clean, removeNumbers)
text_corpus_clean <- tm_map(text_corpus_clean, stripWhitespace)

stop_words <- c(stopwords('english'), "a", "b") 

##Remove more stop words
text_corpus_clean <- tm_map(text_corpus_clean, removeWords, stop_words)

tdm <- TermDocumentMatrix(text_corpus_clean) #or 
dtm <- DocumentTermMatrix(text_corpus_clean, control = list(wordLengths = c(4, Inf)))
inspect(tdm)


#inspect part of the term-document matrix
inspect(tdm[1:10, 1:50])


#Frequent terms that occur between 30 and 50 times in the corpus
frequent_terms <- findFreqTerms (tdm,30,50) 

```

#### WordCloud for Data.World Data
```{r}
#Set seed to get the same wordcloud
set.seed(45)

#install.packages("wordcloud") 
library(wordcloud)

#WordCloud
freq = data.frame(sort(colSums(as.matrix(dtm)), decreasing=TRUE))
wordcloud(rownames(freq), freq[,1], max.words=50, colors=brewer.pal(1, "Dark2"))
```


#### WordCloud 2
```{r}
library(wordcloud)
library(qdap)
#install.packages("qdap")
library(RColorBrewer)
library(RWeka)

BigramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 2, max = 2))
tdm.bigram = TermDocumentMatrix(text_corpus_clean, control = list(tokenize = BigramTokenizer))

##Extract the frequency of each bigram and analyse the twenty most frequent ones.
freq = sort(rowSums(as.matrix(tdm.bigram)), decreasing = TRUE)
freq.df = data.frame(word=names(freq), freq=freq)
head(freq.df, 20)

#visualize the wordcloud   
wordcloud(freq.df$word, freq.df$freq, max.words=100, random.order = T )
```


#### Most Frequent Bigrams
```{r}
#visualize the top 15 bigrams
library(ggplot2)
ggplot(head(freq.df, 15), aes(reorder(word,freq), freq)) +
  geom_bar(stat = "identity") + coord_flip() +
  xlab("Bigrams") + ylab("Frequency") +
  ggtitle("Most Frequent Bigrams")
```


#### What are the top requested highly paid data science skills?
```{r}
#Extract skills and salary column
salary_skills <- jobspikr_data %>%  
  select(inferred_skills, inferred_salary_currency, inferred_salary_time_unit, inferred_salary_from, inferred_salary_to)

#Filter out inferred_salary_currency = 'INR' or 'blank'
salary_skills <- filter(salary_skills, inferred_salary_currency == 'USD')

#Create Inferred_Salary_Median
salary_skills <- mutate(salary_skills, Inferred_Salary_Median = ((inferred_salary_to - inferred_salary_from) / 2) + inferred_salary_from)

#Filter out Inferred_Salary_Median = '0'
salary_skills <- filter(salary_skills, Inferred_Salary_Median > 0)

#Filter out Yearly Salaries
salary_skills_yearly <- filter(salary_skills, inferred_salary_time_unit == 'yearly')
salary_skills_yearly <- mutate(salary_skills_yearly, Inferred_Salary_Median_Annual = Inferred_Salary_Median * 1)
#Drop annual salary = $500 (assume bad data)
salary_skills_yearly <- filter(salary_skills_yearly, Inferred_Salary_Median_Annual > 500)

#Annualize hourly salaries
salary_skills_hourly <- filter(salary_skills, inferred_salary_time_unit == 'hourly')
salary_skills_hourly <- mutate(salary_skills_hourly, Inferred_Salary_Median_Annual = Inferred_Salary_Median * 2080)

#Annualize monthly salaries
salary_skills_monthly <- filter(salary_skills, inferred_salary_time_unit == 'monthly')
salary_skills_monthly <- mutate(salary_skills_monthly, Inferred_Salary_Median_Annual = Inferred_Salary_Median * 12)

#Combine yearly, hourly, & monthly tables
salary_skills <- bind_rows(salary_skills_yearly, salary_skills_hourly, salary_skills_monthly)

#Separate skills from the inferred_skills column into several rows
salary_skills_long <- separate_rows(salary_skills, inferred_skills, sep = "\\|")

#What is the frequency of inferred_skills?
Skills_freq <- count(salary_skills_long, inferred_skills)
Skills_freq <- arrange(Skills_freq, desc(n))

#Merge Inferred_Salary_Median into Skills_freq
distinct_salary_skills <- distinct(salary_skills_long, inferred_skills, .keep_all = TRUE)
distinct_salary_skills <- select(distinct_salary_skills, inferred_skills, Inferred_Salary_Median_Annual)

Skills_freq <- left_join(Skills_freq, distinct_salary_skills, by = 'inferred_skills')
Skills_freq <- mutate(Skills_freq, weighted_sal = n * Inferred_Salary_Median_Annual) #Assume frequency of skill gives weight to Salary

Skills_freq <- arrange(Skills_freq, desc(weighted_sal))
head(Skills_freq, n = 10)
```


#### Plot of the top 20 Highly Paid Data Science Skills
```{r}
Skills_freq_top_sal <- slice(Skills_freq, 1:20)
ggplot(Skills_freq_top_sal, aes(inferred_skills, weighted_sal))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


#### Scrapping data from Indeed Website:
```{r library}
#Load libraries
library(rvest)
library(dplyr)
library(tidyverse)
library(xml2)
library(stringr)

jobs.data<- data.frame(source=character(),
                job.title=character(),
                company.name=character(),
                location=character(),
                job.salary=double(),
                job.description=character())
```


```{r Scrapping-Indeed}
Indeed_data<-data.frame(job.title=character(),
                company.name=character(),
                location=character(),
                job.description=character(),
                result.content=character(),
                job.seen.beacon=character(),
                job.id=character())
pagenumber <- 0
for (i in 0:100) {
    temp = as.character(i*10)
    link <- "https://www.indeed.com/jobs?q=data%20scientist&start=" 
    link <- paste(link,temp,sep="")
    print(link)
    page <- read_html(link)
 
    jobcards<-page %>%html_node(xpath='//*[@id="mosaic-provider-jobcards"]') %>%
                  html_children()
    getdetailjob <- xml_attrs(jobcards)
      
    jobid<-data.frame(job.id=getdetailjob[[1]][1])
                             
    for (j in 2:(length(getdetailjob)-1)){
      tempv <- getdetailjob[[j]][1]
      if (str_detect(tempv,'job_|sj_')) {
        jobid <-rbind(jobid,getdetailjob[[j]][1])
       }  
    }  
    
    #View(jobid)  
   tempdf<-data.frame(job.title=trimws(page %>% 
                            html_nodes(".jobTitle-color-purple > span")  %>% 
                            html_text()),
                     company.name=trimws(page %>% 
                            html_nodes(".companyName")  %>% 
                            html_text()),
                     location=trimws(page %>% 
                            html_nodes(".companyLocation") %>%
                            html_text()),
                     job.description=trimws(page %>% html_nodes(".job-snippet") %>%
                            html_text()),
                     job.seen.beacon=trimws(page %>%html_nodes(".job_seen_beacon")%>%
                            html_text())
                    )
  
   pagenumber = pagenumber + nrow(tempdf)
   print(i)
   tempdf <- cbind(tempdf,jobid)
   Indeed_data <- rbind(Indeed_data, tempdf)
}

Indeed_data<-Indeed_data %>%
  mutate(location.type="")
Indeed_data$location.type= ifelse (str_detect(Indeed_data$location,"Remote")==TRUE,"Remote", "" )
Indeed_data$location<-str_replace(Indeed_data$location,"Remote","")

View(Indeed_data)    
write.csv(Indeed_data,"~/Desktop/Indeed_data.csv")
```


#### Wordcloud for Indeed data
```{r, out.width="100%",fig.align='center'}
library(tidyr)
library(dplyr)
library(textdata)
library(knitr)
library(tidyverse)
library(wordcloud)
library(RColorBrewer)
library(ggplot2)
library(reshape2)
library(ggthemes)
library(png)
library(stringr)

csvfile <- "https://raw.githubusercontent.com/quaere1verum/sps_public/master/data607-001/assignments/project3/csv/Indeed_data.csv"

Indeed <- read.csv(csvfile)

Indeed <-Indeed %>% slice_max(Indeed$job.description, n = 300)

Indeed$job.description<-str_replace_all(Indeed$job.description, "\\n", " and ")
Indeed$job.description<-str_replace_all(Indeed$job.description,"\\.","")
Indeed$job.description<-str_replace_all(Indeed$job.description,":","") 
Indeed$job.description<-str_replace_all(Indeed$job.description,"You will","") 
Indeed$job.description<-str_replace_all(Indeed$job.description,"You'll","") 
Indeed$job.description<-gsub("[…]","",Indeed$job.description)

words<-data.frame(w=character())
words<-strsplit(Indeed$job.description,split = "and|,")

skillset <-data.frame(skill=character())
s <-length(words)

for (i in 1:s){
  for (j in 1:lengths(words[i])){
    rows<-data.frame(skill=str_trim(words[[i]][j]))
    skillset <-rbind(skillset,rows)
  }  
}

word_freq<- skillset %>% group_by(skill)%>%
  summarise(wfreq=n()) 

png("Indeed_wordcloud.png",width = 12, height = 8,units = "in", res=300)
par(mar=rep(0,4))
set.seed(10142021)
word_freq <- word_freq %>% arrange(desc(wfreq))

#set.seed
set.seed(87)
wordcloud(word_freq$skill,freq = word_freq$wfreq,scale=c(3.5, 0.25),
          colors=brewer.pal(8,"Dark2"))

wordcloud_pic <- 'https://github.com/quaere1verum/sps_public/blob/master/data607-001/assignments/project3/Indeed-WordCloud.png'

knitr::include_graphics(wordcloud_pic)
```


#### lm function
```{r}
library(tidyr)
library(dplyr)
library(textdata)
library(knitr)
library(tidyverse)
library(wordcloud)
library(RColorBrewer)
library(ggplot2)
library(reshape2)
library(ggthemes)
library(png)
job_df<- read.csv('https://raw.githubusercontent.com/quaere1verum/sps_public/master/data_scientist_42928_20211010_1633890523765196_1.csv')

test<-strsplit(job_df$inferred_skills, split = "\\|")

skillset <-data.frame(skill=character())
s <-length(test)

for (i in 1:s){
  for (j in 1:lengths(test[i])){
    rows<-data.frame(skill=test[[i]][j])
    skillset <-rbind(skillset,rows)
  }  
}

skill_company <-data.frame(skillid=character(),companyid=character(),state=character())

test <-job_df %>% select (company_name,inferred_skills,state)%>%
    filter(inferred_skills!="")

datarow<-nrow(test)

for (i in 1:datarow){
  infer_byrow <-c(skillid=strsplit(test[[i,2]], split = "\\|")) 
  rows<-data.frame(skillid=infer_byrow,companyid=test[[i,1]],state=test[[i,3]])
  skill_company <-rbind(skill_company,rows)
}

# lm function by state
company_state <- skill_company %>% select(companyid,state)%>%
  group_by(companyid,state) %>%
  summarise(skill_count=n())  

company_state <- company_state %>% mutate(ratio= skill_count/sum(skill_count))

ggplot(company_state %>% filter (state %in% c("CA","NY","TX","NE","FL","NJ","MA")) , aes(sample=skill_count))+
  stat_qq(aes(color =state))+
  stat_qq_line(aes(color = state))+
  facet_grid(~state)

lm_company_state <- lm(ratio~state,data=company_state)
summary(lm_company_state)

lmplot<- company_state %>% filter (state %in% c("CA","NY","TX","NE","FL","NJ","MA"))
        
ggplot(data = lmplot, aes(x = skill_count, y = ratio)) +
  geom_jitter() +
  geom_smooth(method = "lm")+
  facet_grid(~state)+
  xlab("Skills")+
  ylab("Ration")+
  theme(axis.text.x = element_text(size=5),
        axis.text.y = element_text(size=5),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))
```


#### Just one more plot from Data.World Data
```{r}
word_freq<- skillset %>% group_by(skill)%>%
  summarise(wfreq=n())

#Word Frequency  "Values Skills in Data Science
ggplot(top_n(word_freq,35), aes(x=reorder(skill,wfreq),y = wfreq)) + 
  geom_bar(stat='identity',fill="olivedrab")+
  coord_flip()+
  ylab("Count")+
  xlab("Skills")+
  theme_tufte()+
  ggtitle("Valued Skills in Data Science")  
```


#### To answer our initial question "Which are the most valued data science skills?" from our datasets [Data.World](https://data.world/jobspikr/10000-data-scientist-job-postings-from-the-usa/workspace/file?filename=data_scientist_united_states_job_postings_jobspikr.csv] and a downloaded HTML files from Indeed [https://www.indeed.com/jobs?l=New%20York%20State&vjk=ae3c34627b58db77) and from [Indeed.com](https://www.indeed.com/jobs?q=data%20scientist&start=) we were able to conclude that software skills: Python, Machine Learning and SQL rank amongst the highest as well as general skills: analysis, statistics, design, research and management.



